On the Dynamics of Anthropomorphism in Robotics

### 2014-01-21 dynamics-anthropomorphism.tex
Imported changes from 'Anthropomorphism and cognition' HRI workshop

@@

% background shading
\path[fill=gray!20] (0,0) rectangle (0.6,5.5);
\path[fill=gray!50] (0.6,0) rectangle (5.2,5.5);
\path[fill=gray!20] (5.2,0) rectangle (9.8,5.5);
\draw(0,5.5) node[anchor=south west] {\scriptsize \sc Initialization};
\draw(3,5.5) node[anchor=south] {\scriptsize \sc Familiarization};
\draw(7.5,5.5) node[anchor=south] {\scriptsize \sc Stabilization};
% horizontal axis
\draw[->] (0,0) -- (10,0) node[anchor=north] {$t$};
\draw(5,-0.1) node[anchor=north] {\scriptsize Duration of interaction};

@@

\draw (-0.05, 3) -- (0.05, 3) node[anchor=east] {\scriptsize $\text{ICA}_{\mathcal B}$};
\draw (-0.05, 4.1) -- (0.05, 4.1) node[anchor=east] {\scriptsize $\text{ICA}_{\mathcal A}$};
\draw (-0.05, 0.6) -- (0.05, 0.6) node[anchor=east] {\scriptsize $\text{ICA}_{\mathcal C}$};

@@

\caption{Dynamics of anthropomorphism. We distinguish three main phases:
\emph{initialization}, \emph{familiarization} and \emph{stabilization},
preceded by a \emph{pre-interaction} phase. In the pre-interaction phase,
users build an \emph{initial capital of anthropomorphism} (ICA). Once the
interaction starts, the level of anthropomorphism increases due to the
\emph{novelty effect}, and then decreases to reach a \emph{stabilized level
of anthropomorphism} (SLA).  During the interaction, unpredicted behaviors
of the robot (\emph{disruptive behaviors}) may lead to local increase of the
level of anthropomorphism.}

@@

The model of anthropomorphism we propose (Figure~\ref{fig:dynamics}) represents
how the level of anthropomorphic effects (\ie observable manifestations of
anthropomorphism, as defined in section~\ref{sec:intro}) evolves over a
long-term human-robot interaction. By long-term interaction, we mean direct
(non-mediated), repeated interaction with the same robot, over a long period of
time. The nature of the interaction (goal-directed, entertainment, etc.) may
however vary.

Anthropomorphism is quantified by a \emph{normalized level of anthropomorphic
effects}: because anthropomorphic effects are not quantified on an absolute
scale, we present them as a normalized value, that spans from a minimum (no
anthropomorphic effects) to a maximum (corresponding to the novelty effect peak
on Figure~\ref{fig:dynamics}). The actual maximum value of anthropomorphic
effects depends on each unique combinations of human, robot and several other
factors we introduce below, and thus varies. The general \emph{shape} of the
model remains however the same and depicts the evolution of anthropomorphism over
time, \ie the general dynamics of anthropomorphism.

The model takes into account the duration of the interaction, the nature of the
interaction, as well as acquired experience and familiarization mechanisms. We
also formally introduce a so-called \emph{novelty effect} that
models the first phase of human-robot interaction, during which a specific
increase of anthropomorphic interactions is observed.

@@

A robot has an initial potential of being anthropomorphized. It has been shown
that some \textit{people} anthropomorphism more than others, some
\textit{situations} induce anthropomorphism more than others, that
\textit{children} tend to anthropomorphize more than adults, and some
\textit{cultures} are notorious for their anthropomorphic religions and
worldviews \cite{epley_when_2008}. Our model of anthropomorphism takes these
determinants into account and initializes the level of anthropomorphic
interactions between a human and a robot to a value that we call
\textit{initial capital of anthropomorphism} (ICA on
Figure~\ref{fig:dynamics}). The ICA describes the first (real or imagined)
contact to a robot. In this stage of "pre-interaction", people form initial
expectations toward the robot and imagine how they will use / interact with it.

We build the ICA on three main factors that \textit{a priori} determine the
potential that a robot will be anthropomorphized:

@@

    \item \emph{Situation-centered factor}: The real or imagined
        \textbf{purpose} of the robot, including the situational context in
        which it is used. The task context and role in which the robot is used
        / experienced (environmental context)~\cite{joosse_what_2013}.

@@

By \textbf{purpose} we suggest that the real or imagined context in which a
robot is used and the interaction that it brings along, impacts how far the
robot will be attributed human-like characteristics. We draw on findings such
as presented in Joosse \textit{et al.}~\cite{joosse_what_2013}. The authors showed
that when the same robot (NAO) is used in a different task context (cleaning
task \textit{vs.} tour guide), this influences the perception of the
"personality" of the robot. In general, we think that a robot which is imagined
to be used in a social, entertaining or playful context leads to a higher ICA
than a robot which is used for a routine or focused task (security, rescue, etc.). This
idea receives support from \fixme{XX (Kiesler / Goetz?)} work that revealed
that people prefer a serious robot for serious tasks and a less serious robot
for more playful tasks \fixme{REF}. Also, we suggest that the environmental
context in which people experience and interact with the robot impacts the ICA.
When several people who might be friends interact simultaneously with the robot
this might lead to an increased ICA \fixme{study: people are more emotional
when watching TV in company than when watching TV on their own}.

@@

We suggested that the ICA is built on three factors (section~\ref{sec:ica}):
user's \emph{personality}, robot's \emph{design} and interaction \emph{purpose}
(or \emph{interaction context}). The user's personality and the context of use
do also influence the SLA. In particular, it appears that the user's level of
acquaintance with technologies plays an important role in long-term tendency to
anthropomophize (people more familiar with technology understand, and hence
predict, better the behavior of the robot, which in turn lead them more
frequently to ultimately consider the robot as a simple tool).

The robot's design, on the other hand, plays a more subtle role: lasting
anthropomorphic effects have been observed on non-anthropomorphic robots (like
the iRobot Roomba or the military iRobot PackBot\footnote{Rodney Brooks has
reported in keynotes that occasionaly soldiers would give a name to
\emph{their} PackBot and require it to be repared instead of being replaced by
another one in case of incident.}), and on the contrary, anthropomorphic
designs can lead to higher expectation deceptions, resulting in the robot not
being used anymore.

Note that the \emph{Initial Capital of Anthropomorphism} and the
\emph{Stabilized Level of Anthropomorphism} are generally not correlated: one
individual may have high potential of anthropomorphizing (high ICA) at first
sight of a good-looking humanoid robot, and get disappointed by the actual
abilities of the robots, down to routine, non-anthropomorphic, interactions
(low SLA), while another one with the same high ICA may, for instance, create
lasting affective bonds with the same robot, and keep anthropomorphizing it
(higher SLA).

@@

\paragraph{Explanations for anthropomorphism}

According to Lee \textit{et al.} \cite{lee_human_2005},
there are two main perspectives in explaining people's tendency to
anthropomorphize. The first one explains anthropomorphism from the design of
the artifact (anthropomorphic form in the design). It assumes that humans
directly respond to life-like or social cues that an object or system emits,
without thoughtful mental processing, by simply applying stereotypes and
heuristics to it. In fact, from early childhood on, humans are inherently
well-trained to perceive life \cite{epley_seeing_2007}. Schmitz
\cite{schmitz_concepts_2011} describes that within the visual scope of design,
the outer appearance can have an important impact on the overall perception of
an object. The basic assumption here is that if an artifact appears much like a
human, it is likely to be treated similar to a human. If this explanation of
anthropomorphism is correct, people may respond automatically to social cues
emitted by a robot, and apply human-human social schemas and norms to these
interactions.

The second perspective applies a human-centered, cognitive viewpoint where
anthropomorphism is described through people's specific mental model they
construct about how an artifact works the way it does. According to v.
Foerster \fixme{ref} \textit{``we anthropomorphize because it allows us to
explain things we do not understand in terms that we do understand, and what we
understand best is ourselves as human beings"}~\cite{hegel_understanding_2008}.
This is consistent with the \emph{familiarity
thesis}~\cite{hegel_understanding_2008} which claims we understand the world
based upon a mental model of the world that we are most familiar with.
Consequently, people tend to thoughtfully develop a mental model of agents in
their environment and make inferences about it based on what is familiar to
them. This point of view implicitly builds on a person's ability to attribute
mental states to oneself and others (\ie the availability of a \emph{theory of
mind}~\cite{Premack1978a}). A theory of  mind for other agents enables us to
attribute intentionality to those agents
\cite{leslie_pretense_1987,admoni_multi-category_2012}. Previous research
examined the validity of the mental model concept with various kinds of robots
\cite{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people
tend to hold richer mental models about anthropomorphic robots in contrast to
mechanic ones \cite{kiesler_mental_2002}.

@@

\end{tikzpicture}
%}
\caption{The three cognitive phases of anthropomorphism: Phase I is the instinctive,
sub-cognitive identification of living peers. {\it Empathy} is characteristic
of this stage. After longer observation or short, uncontextualized interaction
(typically, a lab environment), the user enters Phase II: the user projects a
mental model he/she is already familiar with onto the robot. After longer {\it
contextualized} interaction (typically, at home), the user enters Phase III of
anthropomorphism: the user recomposes an accurate mental model of the robot,
based on experience. This leads to adapted interaction modalities, that may
still be anthropomorphic, or not.}
\label{fig:cognitivemodel}
\end{figure}

\paragraph{Building cognitive models over time}

Building on these previous researches, Figure~\ref{fig:cognitivemodel} offers a
cognitive perspective on the dynamics of anthropomorphism presented in the
previous section.

The first cognitive phase we identify that contribute to anthropomorphism is
actually \emph{pre-cognitive}. This is supported by studies like Rosenthal-von
der Pütten \textit{et al.}~\cite{rosenthal-vonderputten_experimental_2013} that
investigated the neural correlates of emotional reactions of human towards
robot.

After a longer observation period (typically including complete action
sequences of the robot) or short interaction (touching, short talk like
greetings), we suggest the human enters the cognitive \emph{phase II}: in this
phase, the human starts to build a behavioral and cognitive model of the robot
that would support both the observed and imagined capabilities of the robot.
The \emph{familiarity thesis} would support the idea that
the human first projects onto the robot mental models of similar agents he/she
is already familiar with (ranging from animals to human adults, through
pets and children). We hypothesize that the nature of the projected mental
model, as well as how deep the human engages in this projection, might be
driven by the same parameters as we presented for the \emph{initial capital of
anthropomorphism} (section~\ref{sec:ica}).

We identify a further transition to the cognitive \emph{phase III} after a
\emph{contextualized} interaction. A \emph{contextualized} interaction is
\emph{explicitly purposeful} (the purpose of the interaction, be it purely
entertainment, is explicit and conscious for the human), and takes place in an
environment that fosters a stronger cognitive (and possibly affective)
commitment from the human in the interaction (typically, at home). During this
interaction, the human iteratively restate and reshape its behavioral and
mental model of the robot (\emph{How does the robot react to such and such
situation/input? What does the robot know about me? About itself? About our
environment? What can the robot learn?}, etc.).

This mental process heavily depends on the human understanding of the robot's
inner working, as well as his/her own tendency to anthropomorphize (the
\emph{personality} in ICA factor), but at this stage, the \emph{perception} of
the robot (its shape for instance) and its intended \emph{purpose} play a less
important role. It is mostly a human-centric process.

The result of this third phase is an iteratively adapted cognitive model of the
robot.

\paragraph{Relation to the model of anthropomorphism} It must be noted that the
three cognitive phases we introduce here do not match the
\emph{initialization}, \emph{familiarization} and \emph{stabilization} phases
introduced at~\ref{sec:phases}: in particular, cognitive phases I and II are
both included in the \emph{initialization} phase of the anthropomorphism model.

Sub-cognitive anthropomorphism typically \emph{initiates} the novelty effect by
rapidly engaging the human in the interaction through an initial projected
\emph{agency}, whereas cognitive phase II (projection of familiar mental
models) supports the novelty effect by inducing beliefs that the robot is set
up with possibly complex cognitive abilities.

The cognitive phase II also overlaps with the \emph{familiarization} phase: as
(s)he get used to the robot, we hypothesize the human restates and adapts its
cognitive model of the robot by iteratively reshaping pre-existent, familiar
models until it provides a satisfying support to explain and justify the
observed robot behavior.

A \emph{stable level of anthropomorphism} is reached when the adaptation
process depicted in cognitive phase III reached a stable state, \ie the human
experience with the robot is correctly supported by the cognitive model the
human has built.

@@

@@

Anthropomorphism is traditionally empirically understood as the combination of
the anthropomorphic design of a robot and the psychological determinants of the
user. We have found out that the duration and context of the interaction is a
third factor that plays a key role. In this article, we propose a new formal
model of anthropomorphism that accounts for these three factors and also
explicits the dynamics of anthromorphism. We introduce the concepts of
\emph{initial} and \emph{stabilized levels of anthropomorphism} as compound
factors to characterize the profile of a given anthropomorphic interaction.

The article also discusses the cognitive correlates of anthropomorphism. We
proposes to identify three cognitive phases corresponding to successive
refinements of the mental models of the robot that the user builds during the
interaction. We show how these phases relate to observable anthorpomorphic
behaviours, and how they evolve over time.

While not definitive, we hope that this contribution consolidates the
scientific grounds of anthropomorphism, and provides support for further
research on long-term acceptance of robots in human environments.

>>>>

The phenomenological model of anthropomorphism we propose (Figure~\ref{fig:dynamics},~\cite{lemaignan2014dynamics}) represents
how the level of anthropomorphic effects (\ie observable manifestations of
anthropomorphism, as defined in section~\ref{sec:intro}) evolves over a
long-term human-robot interaction. By long-term interaction, we mean direct
(non-mediated), repeated interaction with the same robot, over a long period of
time.

In this model, anthropomorphism is quantified by a \emph{normalized level of
anthropomorphic effects}: because anthropomorphic effects are not quantified on
an absolute scale, we present them as a normalized value, that spans from a
minimum (no anthropomorphic effects) to a maximum (corresponding to the novelty
effect peak on Figure~\ref{fig:dynamics}). The actual maximum value of
anthropomorphic effects depends on each unique combination of human, robot and
several other factors we introduce below, and thus varies. The general
\emph{shape} of the model remains however the same and depicts the evolution of
anthropomorphism over time, \ie the general dynamics of anthropomorphism.

The model takes into account the duration of the interaction, the nature of the
interaction, as well as acquired experience and familiarization mechanisms. We
also formally introduce a so-called \emph{novelty effect} that models the first
phase of human-robot interaction, during which a specific increase of
anthropomorphic interactions is observed. We focus on \emph{long-term
interaction}, \ie direct (non-mediated), repeated interaction with the same
robot, over an extended period of time (typically longer than a week). The nature 
of the interaction (goal-directed, entertainment, etc.) may
however vary.

@@

% background shading
\path[fill=gray!20] (0,0) rectangle (0.6,5.5);
\path[fill=gray!50] (0.6,0) rectangle (5.2,5.5);
\path[fill=gray!20] (5.2,0) rectangle (9.8,5.5);
\draw(0,5.5) node[anchor=south west] {\scriptsize \sc Initialization};
\draw(4,5.5) node[anchor=south] {\scriptsize \sc Familiarization};
\draw(7.5,5.5) node[anchor=south] {\scriptsize \sc Stabilization};
% horizontal axis
\draw[->] (0,0) -- (10,0) node[anchor=north] {$t$};
\draw(5,-0.1) node[anchor=north] {\scriptsize Duration of interaction};

@@

\draw (-0.05, 3) -- (0.05, 3) node[anchor=east] {\scriptsize ICA$_{\mathcal B}$};
\draw (-0.05, 4.1) -- (0.05, 4.1) node[anchor=east] {\scriptsize ICA$_{\mathcal A}$};
\draw (-0.05, 0.6) -- (0.05, 0.6) node[anchor=east] {\scriptsize ICA$_{\mathcal C}$};

@@

\caption{The dynamics of anthropomorphism. We distinguish three main phases:
\emph{initialization}, \emph{familiarization} and \emph{stabilization},
preceded by a \emph{pre-interaction} phase. In the pre-interaction phase,
users build an \emph{initial capital of anthropomorphism} (ICA). Once the
interaction starts, the level of anthropomorphism increases due to the
\emph{novelty effect}~\cite{kanda_interactive_2004}, and then decreases to reach a \emph{stabilized level
of anthropomorphism} (SLA).  During the interaction, unpredicted behaviors
of the robot (\emph{disruptive behaviors}) may lead to local increase of the
level of anthropomorphism.}

@@

@@

The \emph{initial capital of anthropomorphism} describes the initial potential
for the robot to be anthropomorphized by the human user in a given situation.
This potential depends on several factors. It has been shown, for instance, that
some \textit{people} tend to anthropomorphize more than others, that some
\textit{situations} induce anthropomorphism more than others, that
\textit{children} tend to anthropomorphize more than adults, and that some
\textit{cultures} are notorious for their anthropomorphic religions and
worldviews~\cite{epley_when_2008}. Also the shape and design of the robot play a
role, and the context in which the interaction takes place. Our model of
anthropomorphism takes these determinants into account and initializes the level
of anthropomorphic interactions between a human and a robot to a value that we
call \emph{initial capital of anthropomorphism} (ICA). The ICA describes the
first (real or imagined) contact to a robot. In this stage of pre-interaction,
people form initial expectations toward the robot and imagine how they will use
it / interact with it.

We build the ICA on three main factors that \emph{a priori} determine the
potential that a robot will be anthropomorphized:

@@

    \item \emph{Situation-centered factor}: The real or imagined
        \textbf{purpose} of the robot, including the situational context in
        which it is used, as well as the task context and role in which the robot is
		used / experienced (environmental context)~\cite{joosse_what_2013}.

@@

By taking the \textbf{purpose} of a robot into account, we suggest that the real
or imagined context in which a robot is used and the interaction that this usage
brings along, impacts how far the robot will be attributed human-like
characteristics. We draw on findings such as presented in Joosse
\etal~\cite{joosse_what_2013}. The authors showed for instance that when the
same robot (NAO) is used in a different task context (cleaning task \emph{vs.}
tour guide), users ascribe different ``personalities'' to the robot. In general,
a robot which is imagined to be used in a social, entertaining or playful
context leads to a higher ICA than a robot which is used for a routine or
focused task (security, rescue, etc.). This idea also receives support from
Goetz \& Kiesler's work that revealed that people prefer a serious robot for
serious tasks and a less serious robot for more playful
tasks~\cite{goetz_cooperation_2002, goetz_matching_2003}. Also, we suggest that
the environmental context in which people experience and interact with the robot
impacts the ICA. For instance, several friends interacting simultaneously with
the robot might lead to increased ICAs, due to increased human-human social
interactions (the robot might be perceived to be part of the social interaction,
and in turn attributed human-like qualities)~\cite{baxter2013do}.

@@

We proposed that the ICA is built on three factors: user's \emph{personality},
robot's \emph{design} and interaction \emph{purpose} (or \emph{interaction
context}). The user's personality and the context of use do also influence the
SLA. In particular, it appears that the user's level of acquaintance with
technologies plays an important role in long-term tendency to
anthropomophize~\cite{fink_living_2013} (people more familiar with technology
understand, and hence predict, better the behavior of the robot, which in turn
leads them more frequently to ultimately consider the robot as a simple tool).

The robot's design, on the other hand, plays a more subtle role, and strong
initial anthropomorphic design does not mandate high SLA: lasting
anthropomorphic effects have been observed on non-anthropomorphic robots (like
the iRobot Roomba~\cite{fink_living_2013} or the military iRobot
PackBot\footnote{Rodney Brooks has reported in keynotes that occasionally
soldiers would give a name to \emph{their} PackBot and require it to be repaired
instead of being replaced by another one in case of incident.}), and on the
contrary, anthropomorphic designs can lead to higher expectation deceptions,
resulting in the robot not being used anymore.

Note that the \emph{Initial Capital of Anthropomorphism} and the
\emph{Stabilized Level of Anthropomorphism} are generally not correlated: one
individual may have high potential of anthropomorphizing (high ICA) at first
sight of a good-looking humanoid robot, and get disappointed by the actual
abilities of the robots, down to routine, non-anthropomorphic, interactions
(low SLA), while another user with the same high ICA may, for instance, creates
lasting affective bonds with the same robot, and keeps anthropomorphizing it
(higher SLA).

@@

@@

\end{tikzpicture}
%}
\caption{The three cognitive phases of anthropomorphism: Phase I is the instinctive,
sub-cognitive identification of living peers. {\it Empathy} is characteristic
of this stage. After longer observation or short, non-contextualized interaction
(typically, a lab environment), the user enters Phase II: the user projects a
mental model he/she is already familiar with onto the robot. After longer {\it
contextualized} interaction (typically, at home), the user enters Phase III of
anthropomorphism: the user recomposes an accurate mental model of the robot,
based on experience. This leads to adapted interaction modalities, that may
still be anthropomorphic, or not.}
\label{fig:cognitivemodel}
\end{figure}

We provide in this section a tentative interpretation of anthropomorphism in
terms of cognitive correlates. We warmly welcome the feedback and
comments from both the cognitive sciences and robotics communities to further
discuss these findings.

We propose three different cognitive phases (Figure~\ref{fig:cognitivemodel}),
which do not directly match the previously presented three phases of
anthropomorphism but are still related.

\subsection{Explanations for anthropomorphism}

According to Lee \etal~\cite{lee_human_2005}, there are two main perspectives in
explaining people's tendency to anthropomorphize. First one explains
anthropomorphism from the design of the artifact. It assumes that humans
directly respond to life-like or social cues that an object or system emits,
without thoughtful mental processing, by simply applying stereotypes and
heuristics to it. In fact, from early childhood on, humans are inherently
well-trained to perceive life \cite{epley_seeing_2007}. Schmitz
\cite{schmitz_concepts_2011} describes that within the visual scope of design,
the outer appearance can have an important impact on the overall perception of
an object. The basic assumption here is that if an artifact appears much like a
human, it is likely to be treated similar to a human. If this explanation of
anthropomorphism is correct, people may respond automatically to social cues
emitted by a robot, and apply human-human social schemas and norms to these
interactions.

The second perspective applies a human-centered, cognitive viewpoint where
anthropomorphism is described through people's specific mental model they
construct about how an artifact works the way it does.  We then
anthropomorphize because it allows us to explain things we do not understand in
terms that we do understand, and what we understand best is ourselves as human
beings. This is consistent with the \emph{familiarity
thesis}~\cite{hegel_understanding_2008} which claims that we understand the
world based upon a mental model of the world that we are most familiar with.
Consequently, people tend to thoughtfully develop a mental model of agents in
their environment and make inferences about it based on what is familiar to
them -- humans and human behavior, for instance.  This point of view implicitly
builds on a person's ability to attribute mental states to oneself and others
(\ie the availability of a \emph{theory of mind}~\cite{premack1978does} -- the
link between a tendency to anthropomorphize and the engagement in the
attribution of mental states to other humans has been recently demonstrated at
the brain level in~\cite{cullen2013individual}). A theory of mind for other
agents enables us to attribute intentionality to those agents
\cite{leslie_pretense_1987,admoni_multi-category_2012}. Previous research
examined the validity of the mental model concept with various kinds of robots
\cite{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people
tend to hold richer mental models of anthropomorphic robots in contrast to
mechanic ones \cite{kiesler_mental_2002}.

\subsection{Cognitive Processes and Phases}

The main underlying cognitive process in anthropomorphism is understood as
perceiving and reasoning about something non-human and unfamiliar based on one's
representation of the familiar and well-known concept of being
human~\cite{epley_when_2008}. This led us to interpret the phases of
anthropomorphic interactions as parallel cognitive phases
(Figure~\ref{fig:cognitivemodel}).

The so-called \emph{phase I} is the instinctive, pre-cognitive identification of
living peers. That humans tend to anthropomorphize robots intuitively in this
pre-cognitive way
is supported by studies done by Rosenthal-von der Pütten
\textit{et al.}~\cite{rosenthal-vonderputten_experimental_2013} who investigated
the neural correlates of emotional reactions of humans towards a robot. {\it
Empathy} is characteristic of this stage~\cite{rosenthalvonderPutten2013neural}.
Anthropomorphism at this pre-cognitive stage might also be mediated by the
human's mirror neurons system (neurons that fire both during execution of
specific goal-oriented action and during the viewing of action directed toward
the same goal~\cite{Rizzolatti1996, Kilner2009}) by allowing for a mapping of
the robot's goal-directed actions into the human own motor
repertoire~\cite{Gallese1998, Wolpert2003, cullen2013individual}.

After a longer observation period (typically including complete action sequences
of the robot) or short interaction (touching, short talk like greetings), we
suggest the human enters the cognitive \emph{phase II}: in this phase, the human
starts building a behavioral and cognitive model of the robot that would support
both the observed and imagined capabilities of the robot.  The \emph{familiarity
thesis}~\cite{hegel_understanding_2008} supports the idea that the human
first projects onto the robot mental models of similar agents he/she is already
familiar with (ranging from animals to human adults, to pets and children). We 
hypothesize that the nature of the projected mental
model, as well as how deep the human engages in this projection, might be
driven by the same parameters as we mentioned for the \emph{initial capital of
anthropomorphism} (section~\ref{sec:ica}).

The cognitive \emph{phase III} occurs after a \emph{contextualized} interaction.
A \emph{contextualized} interaction is \emph{explicitly purposeful} (the purpose
of the interaction, be it purely entertainment, is explicit and conscious to the
human), and takes place in an environment that fosters a stronger cognitive (and
possibly affective/social) commitment from the human in the interaction
(typically, at home). During this interaction, the human iteratively restates
and reshapes his/her behavioral and mental model of the robot (\emph{How does
the robot react to such and such situation/input?  What does the robot know
about me? About itself? About our environment? What can the robot learn?}, etc.).

This mental process depends on the human understanding of the robot's
inner working, as well as his/her own tendency to anthropomorphize (the
\emph{personality} in ICA factor), but at this
stage, the \emph{perception} of the robot (its shape for instance) and its
intended \emph{purpose} play a less important role. It is mostly a human-centric
process.  The result of this third phase would be an iteratively adapted
cognitive model of the robot.

\subsection{Relation to the model of anthropomorphism}

These cognitive phases overlap but do not exactly match the
\emph{Initialization}, \emph{Familiarization} and \emph{Stabilization} phases
introduced in our model of the dynamics of anthropomorphism. In particular,
cognitive phases I and II are both included in the \emph{initialization} phase
of the anthropomorphism model. Sub-cognitive anthropomorphism typically
\emph{initiates} the novelty effect by rapidly engaging the human in the
interaction through an initial projected agency, whereas cognitive phase II
(projection of familiar mental models) supports the novelty effect by inducing
beliefs that the robot is set up with possibly complex cognitive abilities.

The cognitive phase III also overlaps with the \emph{familiarization} phase: as
the human gets used to the robot, we hypothesize one restates and adapts its
cognitive model of the robot by iteratively reshaping pre-existent, familiar
models until it provides a satisfying support to explain and justify the
observed robot behavior.

A \emph{stable level of anthropomorphism} is reached when the adaptation process
depicted in cognitive phase III reached a stable state, \ie the user's experience
with the robot is correctly supported by the cognitive model he/she has
built.

@@

\section{Discussion}

This interpretation of anthropomorphism in terms of cognitive correlates
opens questions that we hope could be fruitfully discussed during the workshop.
By adopting both a cognitive and a dynamic perspective on the anthropomorphic
bonds that establish between a robot and a human during an interaction, 
some observations can be raised.

For instance, we propose that the human \emph{adapts} iteratively to the robot
by restating its cognitive model of the robot. Could our two models be
conversely relied on to have the robot itself iteratively adapting its
behaviour? In particular, we hypothesize that the so-called novelty effect
represents a peak in anthropomorphic manifestations, followed by mostly
deception-driven refinements of the cognitive behavioural model of the robot.
Could we imagine that the robot pro-actively \emph{verbalises} its own limits,
in a timely manner (likely before the end of the novely effect). And how would this
affect the cognitive models built by the human?

One related question we plan to elaborate on is the effects of \emph{mutual
explicitation of the mental model of the other agent}: a human and a robot
interact on a given task, and after a while, we interrupt the task and ask each
of the agents to explicit the mental model it has built of its partner
(emotional state, beliefs, intentions, etc.). Depending on the cognitive phase
the two agents are in, we may expect varying accuracy levels in the produced
mental models.

Another question raised by these models relates to their transposition to
human-human interaction: while concepts like \emph{novelty effect} are not
directly meaningful when analyzing human-human interaction, insights from
cognitive sciences (and in particular social psychology) regarding how we, as
humans, cope with unexpected behaviours from peers, or on the dynamics of
mental model refinements, could bring interesting perspectives with possible
applications to better manage long-term human-robot interactions.

@@

While anthropomorphism is traditionally empirically understood as the interactions between
the anthropomorphic design of a robot and the psychological determinants of the
user, it appears that the duration and context of the
interaction are also key factors, and that anthropomorphism needs to be
understood as a dynamic phenomenon. In this article, we propose a new formal
model of anthropomorphism that accounts for these three factors and also
explicits the dynamics of anthromorphism. We introduce the concepts of
\emph{initial} and \emph{stabilized levels of anthropomorphism} as compound
factors to characterize the profile of a given anthropomorphic interaction.

The article also discusses the cognitive correlates of anthropomorphism. We
proposes to identify three cognitive phases corresponding to successive
refinements of the mental models of the robot that the user builds during the
interaction. We show how these phases relate to observable anthropomorphic
effects, and how they evolve over time.

While subject to discussion and further extensions, we hope that this
contribution consolidates the scientific grounds of anthropomorphism, and
provides support for a better understanding of long-term acceptance of robots
in human environments.
